\documentclass[11pt]{article}
\usepackage[a4paper,margin=1in]{geometry}
\usepackage{amsmath,amssymb,mathtools,bm}
\usepackage{hyperref}
\usepackage{physics}
\usepackage{enumitem}

\newcommand{\Mpl}{M_{\mathrm{pl}}}

\title{Checking the Mode Equations in\\ \textit{Cosmological gravitational particle production of massive spin-2 particles}}
\author{Project notes for \texttt{Spin2CGPP}}
\date{\today}

\begin{document}
\maketitle

\section{Scope}
These notes document a reproducible check of the mode equations in Siyang Ling, Edward W. Kolb, Andrew J. Long, and Rachel A. Rosen, \texttt{arXiv:2302.04390}. The target equations are the sector-by-sector equations of motion in section 3 of the paper:
\begin{itemize}[leftmargin=1.5em]
\item minimal matter coupling: tensor, vector, and canonical scalar equations;
\item nonminimal matter coupling: tensor, vector, and scalar equations.
\end{itemize}

The scripts in \texttt{Code/script/} start from the published quadratic sector Lagrangians and reproduce the final mode equations by explicit Euler-Lagrange variation, including the canonical field redefinitions used in the paper.

\section{End-to-end status}
The automated checks now operate at two levels. First, \texttt{Code/script/derive\_eoms.wls} verifies the published sector Lagrangians against the quoted mode equations for the minimal tensor, vector, and canonical scalar systems and for the nonminimal tensor, vector, and scalar systems. Second, the covariant-action checks go behind the published reduced Lagrangians where this has been implemented. The minimal tensor diagnostic \texttt{Code/script/check\_covariant\_tensor\_sector.wls} reduces the covariant Fierz--Pauli action to the tensor sector up to an explicit boundary term and verifies the corresponding mode equation. The minimal scalar scripts reduce the covariant scalar action through the constraint chain and match the paper's reduced coefficient basis.

With the current scripts, the original end-to-end goal is reached for the minimally coupled scalar sector, which was the nontrivial coupled system, and for the minimally coupled tensor sector. The vector checks and the nonminimal tensor/vector/scalar checks reproduce the mode equations from the published reduced sector Lagrangians. The remaining open extension is to automate the nonminimal scalar reduction directly from its covariant action.

\section{What the automation does}
The derivation pipeline is organized around a single repeated pattern. If a sector Lagrangian has the form
\begin{equation}
\mathcal{L} = K(\eta)\, q'^{\,2} - M(\eta)\, q^2,
\end{equation}
then after the canonical redefinition
\begin{equation}
q(\eta) = \frac{\chi(\eta)}{\sqrt{2K(\eta)}}
\end{equation}
and after discarding a total derivative, the Lagrangian becomes
\begin{equation}
\mathcal{L} = \frac{1}{2}\chi'^{\,2} - \frac{1}{2}\omega_k^2(\eta)\chi^2,
\qquad
\omega_k^2(\eta) = \frac{4KM + (K')^2 - 2K K''}{4K^2}.
\end{equation}
The Mathematica helper \texttt{CanonicalOmegaSquared} implements exactly this identity.

For the minimal scalar sector, the paper already presents the canonically normalized two-field system
\begin{equation}
\mathcal{L}_S =
\frac12 |\chi_\Pi'|^2 - \frac12 \omega_\Pi^2 |\chi_\Pi|^2
+ \frac12 |\chi_{\mathcal B}'|^2 - \frac12 \omega_{\mathcal B}^2 |\chi_{\mathcal B}|^2
+ \sigma_1 \chi_\Pi^\ast \chi_{\mathcal B}' - \sigma_0 \chi_\Pi^\ast \chi_{\mathcal B}.
\end{equation}
In the Mathematica checks, I model the derivative mixing with an equivalent real-field convention. That convention differs from the paper's complex-mode presentation by a total-derivative reshuffling, so an explicit $\sigma_1'/2$ shift appears in the off-diagonal mass-like term. After accounting for that convention choice, the coupled equations agree with the structure quoted in the paper:
\begin{align}
\chi_\Pi'' + \omega_\Pi^2 \chi_\Pi - \sigma_1 \chi_{\mathcal B}' + \sigma_0 \chi_{\mathcal B} &= 0, \\
\chi_{\mathcal B}'' + \omega_{\mathcal B}^2 \chi_{\mathcal B} + \sigma_1 \chi_\Pi' + \sigma_0 \chi_\Pi &= 0.
\end{align}

\section{Minimal matter coupling}
\subsection{Tensor sector}
The paper starts from
\begin{equation}
\mathcal{L}_T = \frac12 a^2 \left[ D_{ij}' D_{ij}' - \partial_k D_{ij}\partial_k D_{ij} - a^2 m^2 D_{ij}D_{ij}\right].
\end{equation}
After Fourier transforming and replacing $\partial_k \partial_k \to -k^2$, one obtains a single-mode Lagrangian of the form
\begin{equation}
\mathcal{L}_{T,k} = \frac12 a^2 \left[ D'^2 - (k^2 + a^2 m^2) D^2 \right].
\end{equation}
Using $\chi = aD$, the code shows
\begin{equation}
\mathcal{L}_{T,k} = \frac12 \chi'^2 - \frac12 \left(k^2 + a^2 m^2 - \frac{a''}{a}\right)\chi^2
\end{equation}
up to a total derivative. With $H = a'/a^2$, the background identity
\begin{equation}
\frac{a''}{a} = 2a^2 H^2 + aH'
\end{equation}
gives
\begin{equation}
\omega_k^2 = k^2 + a^2 m^2 - 2a^2 H^2 - aH',
\end{equation}
which matches eq.~(3.8) of the paper.

The script \texttt{Code/script/check\_covariant\_tensor\_sector.wls} performs the same check one step earlier, starting from the covariant lower-index tensor field. The reduced covariant expression differs from the sector Lagrangian by the total derivative
\begin{equation}
\frac{d}{d\eta}\left(a^3 H D^2\right),
\end{equation}
and the script verifies that the sector difference after removing this boundary term is zero. It also verifies the Euler--Lagrange equation
\begin{equation}
D'' + 2aH D' + (k^2+a^2m^2)D = 0,
\end{equation}
which is equivalent to the canonical equation above after $\chi=aD$.

\subsection{Vector sector}
The reduced Fourier-space Lagrangian quoted in the paper is
\begin{equation}
\mathcal{L}_{V,k} = K_C |C'|^2 - M_C |C|^2,
\qquad
K_C = \frac{a^4 k^2 m^2}{k^2 + a^2 m^2},
\qquad
M_C = a^4 k^2 m^2.
\end{equation}
Applying the generic canonical-normalization identity above gives
\begin{equation}
\omega_k^2 =
\frac{4K_C M_C + (K_C')^2 - 2 K_C K_C''}{4K_C^2}.
\end{equation}
The script rewrites this result as
\begin{equation}
\omega_k^2 = k^2 + a^2 m^2 - \frac{f''}{f},
\qquad
f = \frac{a^2}{\sqrt{k^2 + a^2 m^2}},
\end{equation}
which is exactly eq.~(3.13).

\subsection{Scalar sector}
The minimally coupled scalar sector is the technically hardest part of the paper. The current project now reconstructs the scalar reduction directly from the covariant quadratic action. The scripts
\begin{itemize}[leftmargin=1.5em]
\item \texttt{Code/script/scalar\_equation\_rank\_check.wls},
\item \texttt{Code/script/derive\_scalar\_mode\_eoms.wls}, and
\item \texttt{Code/test/test\_scalar\_mode\_chain.wls}
\end{itemize}
implement the scalar SVT ansatz, perform the field redefinition
\begin{equation}
\hat{\varphi}_v = \varphi_v - \frac{a^{-1}\bar\phi'}{\Mpl H} A,
\end{equation}
and then eliminate the constrained variables in the order
\begin{equation}
F_\pm \longrightarrow A_\pm \longrightarrow E_\pm.
\end{equation}

This ordering is useful for checking the equations, but the reduced action requires a slightly different bookkeeping step. The fields $E_\pm$ and $F_\pm$ do not carry quadratic kinetic terms, but the cleaned covariant Lagrangian still contains their first derivatives. Therefore \texttt{derive\_scalar\_reduced\_action.wls} first integrates those derivatives by parts, making $E_\pm$ and $F_\pm$ algebraic, and only then solves their constraint equations. After this action-level elimination the apparent $A_+'A_-'$ coefficient cancels exactly. Integrating the remaining $A_\pm'$ terms by parts makes $A_\pm$ algebraic, so it can also be eliminated at the level of the reduced action.

The resulting equations of motion are denoted \texttt{finalB} and \texttt{finalP} in \texttt{derive\_scalar\_mode\_eoms.wls}. The test script verifies three structural facts directly:
\begin{enumerate}[leftmargin=1.5em]
\item both final equations contain $B''$ and $\hat\varphi_v''$;
\item the final $2\times2$ acceleration matrix has rank $2$;
\item the derivation starts from the covariant action rather than the published reduced scalar action.
\end{enumerate}

The current project now includes an explicit Mathematica transcription of the full published scalar coefficient set $K_\varphi$, $M_\varphi$, $K_B$, $M_B$, $L_2$, $L_1$, and $L_0$, together with the derived combinations $\kappa$ and $K_\mathcal{B}$, in \texttt{Code/script/paper\_scalar\_coeffs.wl}. The script \texttt{Code/script/derive\_scalar\_reduced\_action.wls} extracts the same coefficients from the covariant reduced action. It uses the paper's integration-by-parts convention: diagonal one-derivative terms such as $\hat\varphi_v'\hat\varphi_v^\ast$ and $B'B^\ast$ are moved into $M_\varphi$ and $M_B$, while the mixed term $\hat\varphi_v'B^\ast$ is moved into $L_1$ and $L_0$.

The regression check \texttt{Code/test/test\_scalar\_paper\_coeffs.wls} verifies that the residuals for
\begin{equation}
K_\varphi,\quad M_\varphi,\quad K_B,\quad M_B,\quad L_2,\quad L_1,\quad L_0
\end{equation}
vanish symbolically. It also checks both Fourier orientations for the mixed coefficients.

\subsubsection{Scalar-sector symbolic record}
This subsection records the explicit intermediate expressions used by the Mathematica derivation of the minimally coupled scalar equations. The scalar ansatz is evaluated for a pair of complex Fourier modes $u=e^{ikz}$ and $u^{-1}=e^{-ikz}$:
\begin{align}
v_{00} &= a^2 (E_+u+E_-u^{-1}), &
v_{03} &= a^2 \partial_z(F_+u+F_-u^{-1}), \\
v_{11} &= a^2 (A_+u+A_-u^{-1}), &
v_{22} &= a^2 (A_+u+A_-u^{-1}), \\
v_{33} &= a^2 \left[A_+u+A_-u^{-1}+\partial_z^2(B_+u+B_-u^{-1})\right], &
\varphi_v &= \varphi_+u+\varphi_-u^{-1}.
\end{align}
The field redefinition used before solving constraints is
\begin{equation}
\varphi_\pm = \hat\varphi_\pm + \frac{\bar\phi'}{\Mpl aH} A_\pm .
\end{equation}
After the redefinition, every expression is reduced with
\begin{align}
a' &= a^2H, &
a'' &= 2a^3H^2+a^2H', &
H' &= -\frac{\bar\phi'^2}{2a\Mpl^2}, \\
\bar\phi'' &= -2aH\bar\phi' - a^2V_{,\phi}, &
V_{,\phi}' &= V_{,\phi\phi}\bar\phi'.
\end{align}
The compact mass combination and common scalar denominator are
\begin{align}
m_H^2 &\equiv 2H^2-\frac{\bar\phi'^2}{a^2\Mpl^2}, \\
\Delta &\equiv H^2k^4
3a^2(m^2-m_H^2)H^2k^2
\frac{3}{8}a^4m^2\left(6m^2H^2-4H^2m_H^2-m_H^4\right).
\end{align}

At the action level, the first-derivative terms in $E_\pm$ and $F_\pm$ are integrated by parts before imposing their algebraic equations. This gives the diagnostic result
\begin{equation}
\frac{\partial^2\mathcal L}{\partial A_+'\partial A_-'}\bigg|_{E,F\ \mathrm{eliminated}} = 0,
\end{equation}
so $A_\pm$ can also be made algebraic by integration by parts. Eliminating $A_\pm$ produces the reduced scalar action
\begin{align}
\mathcal L_{S,\mathrm{red}}={}&
2K_\varphi\,\hat\varphi_+'\hat\varphi_-'
-2M_\varphi\,\hat\varphi_+\hat\varphi_-
2K_B\,B_+'B_-'
-2M_B\,B_+B_- \nonumber \\
&+L_2\left(\hat\varphi_+'B_-'+\hat\varphi_-'B_+'\right)
+L_1\left(\hat\varphi_+B_-'+\hat\varphi_-B_+'\right)
-L_0\left(\hat\varphi_+B_-+\hat\varphi_-B_+\right),
\end{align}
up to total derivatives and the same overall Fourier normalization used in the scripts. The extracted coefficients are
\begin{align}
K_\varphi &=
\frac{a^2}{2\Delta}
\left[
H^2k^4+3a^2(m^2-m_H^2)H^2k^2
\frac94 a^4m^2(m^2-m_H^2)H^2
\right], \\
K_B &=
\frac{a^6m^2k^4}{8\Delta}
\left(8m^2H^2-6H^2m_H^2-m^2m_H^2\right), \\
L_2 &=
\frac{a^3m^2\bar\phi'}{2\Mpl H\,\Delta}
\left[
H^2k^4+\frac32 a^2(m^2-m_H^2)H^2k^2
\right], \\
L_1 &=
-\frac{a^4m^2\bar\phi'}{\Mpl\,\Delta}
\left[
\left(H^2-\frac14m_H^2-\frac{aHV_{,\phi}}{2\bar\phi'}\right)k^4
-\frac32a^2(m^2-m_H^2)
\left(H^2+\frac14m_H^2+\frac{aHV_{,\phi}}{2\bar\phi'}\right)k^2
\right], \\
L_0 &=
\frac{a^3m^2\bar\phi'}{2\Mpl H\,\Delta^2}
\left(c_{10}^{(L)}k^{10}+c_8^{(L)}k^8+c_6^{(L)}k^6+c_4^{(L)}k^4+c_2^{(L)}k^2\right), \\
M_\varphi &=
\frac{a^2}{2\Delta^2}
\left(c_{10}^{(\varphi)}k^{10}+c_8^{(\varphi)}k^8+c_6^{(\varphi)}k^6
+c_4^{(\varphi)}k^4+c_2^{(\varphi)}k^2+c_0^{(\varphi)}\right), \\
M_B &=
\frac{a^6m^2}{8\Delta^2}
\left(c_{10}^{(B)}k^{10}+c_8^{(B)}k^8+c_6^{(B)}k^6+c_4^{(B)}k^4\right).
\end{align}
The two derived combinations used in the subsequent canonical scalar variables are
\begin{equation}
\kappa=-\frac{L_2}{2k^2K_\varphi},
\qquad
K_{\mathcal B}=\frac{4K_\varphi K_B-L_2^2}{4k^4K_\varphi}.
\end{equation}
The polynomial coefficients in these expressions are
\begin{align}
c_{10}^{(\varphi)} &= H^4, \\
c_8^{(\varphi)} &=
\frac12a^2H^2\left(
12m^2H^2+8H^4-14H^2m_H^2-m_H^4
+\frac{4HV_{,\phi}\bar\phi'}{a\Mpl^2}+2H^2V_{,\phi\phi}
\right), \\
c_6^{(\varphi)} &=
\frac38a^4H^2\bigg[
36m^4H^2+72m^2H^4-82m^2H^2m_H^2-64H^4m_H^2
-7m^2m_H^4+40H^2m_H^4+8m_H^6 \nonumber\\
&\hspace{7em}
+\frac{8(3m^2-4m_H^2)HV_{,\phi}\bar\phi'}{a\Mpl^2}
+16(m^2-m_H^2)H^2V_{,\phi\phi}
\bigg], \\
c_4^{(\varphi)} &=
\frac38a^6\bigg[
4H^2\left(
9m^6H^2+36m^4H^4+16m^2H^6-30m^4H^2m_H^2-76m^2H^4m_H^2 \right. \nonumber\\
&\hspace{8em}\left.
-3m^4m_H^4+31m^2H^2m_H^4+24H^4m_H^4+6m^2m_H^6-6H^2m_H^6-3m_H^8
\right) \nonumber\\
&\hspace{7em}
-\frac{4m^2H^2(H^2-m_H^2)V_{,\phi}^2}{\Mpl^2} \nonumber\\
&\hspace{7em}
+\left(36m^4H^2+8m^2H^4-94m^2H^2m_H^2+m^2m_H^4+48H^2m_H^4\right)
\frac{HV_{,\phi}\bar\phi'}{a\Mpl^2} \nonumber\\
&\hspace{7em}
+\left(36m^4H^2-58m^2H^2m_H^2-m^2m_H^4+24H^2m_H^4\right)H^2V_{,\phi\phi}
\bigg], \\
c_2^{(\varphi)} &=
\frac{9}{32}a^8m^2\bigg[
H^2\left(
18m^6H^2+120m^4H^4+128m^2H^6-78m^4H^2m_H^2-384m^2H^4m_H^2 \right. \nonumber\\
&\hspace{8em}\left.
-9m^4m_H^4+132m^2H^2m_H^4+128H^4m_H^4+23m^2m_H^6-32H^2m_H^6-16m_H^8
\right) \nonumber\\
&\hspace{7em}
-\frac{8H^2(2m^2H^2-2m^2m_H^2+m_H^4)V_{,\phi}^2}{\Mpl^2} \nonumber\\
&\hspace{7em}
+4\left(6m^4H^2-22m^2H^2m_H^2+m^2m_H^4+14H^2m_H^4\right)
\frac{HV_{,\phi}\bar\phi'}{a\Mpl^2} \nonumber\\
&\hspace{7em}
+4(m^2-m_H^2)(12m^2H^2-10H^2m_H^2-m_H^4)H^2V_{,\phi\phi}
\bigg], \\
c_0^{(\varphi)} &=
\frac{27}{32}a^{10}m^4\bigg[
-\frac{2H^2(2m^2H^2-2m^2m_H^2+m_H^4)V_{,\phi}^2}{\Mpl^2}
-\frac{m^2(2H^2-m_H^2)(4H^2+m_H^2)HV_{,\phi}\bar\phi'}{a\Mpl^2} \nonumber\\
&\hspace{7em}
+(m^2-m_H^2)(6m^2H^2-4H^2m_H^2-m_H^4)H^2V_{,\phi\phi}
\bigg].
\end{align}
\begin{align}
c_{10}^{(B)} &= H^2(8m^2H^2-8H^4-2H^2m_H^2-m^2m_H^2), \\
c_8^{(B)} &=
a^2H^2\left[
30m^4H^2+32m^2H^4-96H^6-3m^4m_H^2-56m^2H^2m_H^2+48H^4m_H^2 \right. \nonumber\\
&\hspace{8em}\left.
+5m^2m_H^4+6H^2m_H^4
+(4m^2-24H^2)\frac{HV_{,\phi}\bar\phi'}{a\Mpl^2}
\right], \\
c_6^{(B)} &=
\frac38a^4m^2\left[
96m^4H^4+144m^2H^6-6m^4H^2m_H^2-252m^2H^4m_H^2-192H^6m_H^2 \right. \nonumber\\
&\hspace{8em}\left.
+8m^2H^2m_H^4+200H^4m_H^4-10H^2m_H^6-m^2m_H^6
+(8m^2m_H^2-16H^2m_H^2)\frac{HV_{,\phi}\bar\phi'}{a\Mpl^2}
\right], \\
c_4^{(B)} &=
\frac38a^6m^4\left[
36m^4H^4-48m^2H^6+64H^8-12m^2H^4m_H^2-32H^6m_H^2
-12m^2H^2m_H^4 \right. \nonumber\\
&\hspace{8em}\left.
+4H^4m_H^4+12H^2m_H^6-3m^2m_H^6+2m_H^8 \right. \nonumber\\
&\hspace{8em}\left.
-(24m^2H^2-16H^4-12m^2m_H^2-8H^2m_H^2+8m_H^4)
\frac{HV_{,\phi}\bar\phi'}{a\Mpl^2}
\right].
\end{align}
\begin{align}
c_{10}^{(L)} &= H^4, \\
c_8^{(L)} &=
\frac12a^2H^4\left(9m^2+12H^2-13m_H^2-\frac{4aHV_{,\phi}}{\bar\phi'}\right), \\
c_6^{(L)} &=
\frac38a^4H^2\left[
18m^4H^2+32m^2H^4+64H^6-48m^2H^2m_H^2-64H^4m_H^2
+m^2m_H^4+28H^2m_H^4 \right. \nonumber\\
&\hspace{8em}\left.
+8(-4m^2H^2+4H^4+m^2m_H^2)\frac{aHV_{,\phi}}{\bar\phi'}
\right], \\
c_4^{(L)} &=
\frac{3}{16}a^6m^2H^2\left[
18m^4H^2-24m^2H^4+256H^6-54m^2H^2m_H^2-160H^4m_H^2 \right. \nonumber\\
&\hspace{8em}\left.
+9m^2m_H^4+60H^2m_H^4-7m_H^6
+4(-30m^2H^2+32H^4+12m^2m_H^2+4H^2m_H^2-7m_H^4)
\frac{aHV_{,\phi}}{\bar\phi'}
\right], \\
c_2^{(L)} &=
\frac{9}{16}a^8m^4H^2(2H^2-m_H^2)
\left[
-(4H^2+m_H^2)(3m^2-4H^2-m_H^2)
+4(-3m^2+2H^2+2m_H^2)\frac{aHV_{,\phi}}{\bar\phi'}
\right].
\end{align}

The symbolic comparison performed in \texttt{check\_scalar\_paper\_coeffs.wls} returns
\begin{equation}
\begin{gathered}
\frac{\partial^2\mathcal L}{\partial A_+'\partial A_-'}=0, \\
\delta K_\varphi=\delta M_\varphi=\delta K_B=\delta M_B
=\delta L_2=\delta L_1=\delta L_0=0,
\end{gathered}
\end{equation}
with the same zero residuals in the opposite Fourier orientation. The numerical spot check used by the script,
\begin{equation}
a=2,\quad H=\frac{3}{10},\quad \bar\phi'=\frac15,\quad \Mpl=5,\quad
m=\frac7{10},\quad k=\frac{11}{10},\quad
V_{,\phi}=\frac{13}{10},\quad V_{,\phi\phi}=\frac{17}{10},
\end{equation}
gives
\begin{align}
K_\varphi &= 1.9995900442618875, &
M_\varphi &= 16.623735495848287, &
K_B &= 0.934701933108572, \\
M_B &= 3.146452918424948, &
L_2 &= 0.08485818996827352, &
L_1 &= 2.2787374727884644, &
L_0 &= -0.9584455789624143.
\end{align}

\section{Nonminimal matter coupling}
\subsection{Tensor sector}
The tensor-sector action in the nonminimal model is
\begin{equation}
\mathcal{L}_T = \frac12 a^2 \left[D'^2 - \left(k^2 + a^2(m^2-\Lambda+3H^2+2a^{-1}H')\right)D^2\right].
\end{equation}
Again using $\chi = aD$, the canonical form is
\begin{equation}
\mathcal{L}_{T,k} = \frac12 \chi'^2 - \frac12 \omega_k^2 \chi^2
\end{equation}
with
\begin{equation}
\omega_k^2 = k^2 + a^2 m^2 - a^2 \Lambda + a^2 H^2 + aH',
\end{equation}
matching eq.~(3.58).

\subsection{Vector and scalar sectors}
The paper reduces both the vector and scalar sectors to single-field quadratic actions of the same form
\begin{equation}
\mathcal{L} = K(\eta)\,q'^{\,2} - M(\eta)\,q^2.
\end{equation}
The Mathematica scripts apply the same canonical-normalization identity and reproduce the quoted frequency formulas
\begin{align}
\omega_{V,k}^2 &= \frac{4K_C M_C + (K_C')^2 - 2K_C K_C''}{4K_C^2}, \\
\omega_{S,k}^2 &= \frac{4K_B M_B + (K_B')^2 - 2K_B K_B''}{4K_B^2},
\end{align}
which are eqs.~(3.66) and (3.73).

\section{Files}
\begin{itemize}[leftmargin=1.5em]
\item \texttt{Code/script/Spin2CGPP.wl}: derivation helpers and symbolic checks.
\item \texttt{Code/script/derive\_eoms.wls}: human-readable driver that prints each check.
\item \texttt{Code/script/check\_covariant\_tensor\_sector.wls}: covariant minimal tensor-sector reduction, boundary-term check, and tensor mode equation check.
\item \texttt{Code/script/derive\_scalar\_mode\_eoms.wls}: covariant-to-scalar sequential elimination chain for the minimally coupled scalar sector.
\item \texttt{Code/script/derive\_scalar\_reduced\_action.wls}: action-level scalar reduction and extraction of the paper coefficient basis.
\item \texttt{Code/script/check\_scalar\_paper\_coeffs.wls}: diagnostic driver that prints the full coefficient comparison.
\item \texttt{Code/script/check\_ds\_scalar\_limit.wls}: de Sitter-limit diagnostic for the pure spin-2 scalar reduction.
\item \texttt{Code/test/run\_tests.wls}: Mathematica regression tests.
\item \texttt{Code/test/test\_scalar\_mode\_chain.wls}: regression test for the covariant scalar elimination chain.
\item \texttt{Code/test/test\_scalar\_paper\_coeffs.wls}: regression test for the full scalar coefficient match.
\item \texttt{Code/test/run\_tests.sh}: shell entry point for the tests.
\end{itemize}

\section{Current limitation}
The minimally coupled scalar-sector derivation is now automated from the covariant quadratic action through the paper's reduced-action coefficient basis and the coupled mode-equation structure. The remaining limitation is separate: the nonminimal scalar-sector derivation from the covariant action has not yet been automated at the same level.

\end{document}
